### 2024-11-26 paper/paper.qmd
setup basic structure for Data section & added content points to discussion

---
title: "My title"
subtitle: "Impact Analysis of Japanese Occupation on Population Shifts Across Shanghai’s Districts During WWII"
author: 
  - Sandy Yu
thanks: "Code and data are available at: https://github.com/Jingying-yu/Shanghai_population_change"
date: today
date-format: long
abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
format: pdf
number-sections: true
bibliography: references.bib
---

```{r}
#| include: false
#| warning: false
#| message: false

library(tidyverse)
library(palmerpenguins)
```

# Introduction

Overview paragraph

Estimand paragraph

Results paragraph

Why it matters paragraph

Telegraphing paragraph: The remainder of this paper is structured as follows. @sec-data....

# Data {#sec-data}

## Overview

We use the statistical programming language R [@citeR].... Our data [@shelter].... Following @tellingstories, we consider...

## Data Sources
- talk about where the data in my primary reference book is gathered
- mentions the credibility of source and unstability of recording instrument & methodology

## Historical Background
Place: Shanghai, China
When: 1936-1942
Who: Chinese population in Shanghai

Define:
1. give 1-2 sentence broad overview of China's state of unrest
2.THREE districts in Shanghai: Chinese District, International Settlement, French Concession
- who controlled each district and the level of governance each authority have in comparison to Chinese government
3.Outline area (%) of each district (do not get into specifics, put that in Results section)

Important Event Timeline
1. 1937-08-13: Japanese armed forces entered Shanghai
2. 1937-11-12: Japanese armed forces claims occupation of Shanghai --> ends Chinese district 
3. 1942-01: Japanese armed forces claims authority over International Settlement (which was mainly under the governance of U.K and U.S prior to this date)
4. 1945-08-15: Japan surrendered in WWII
5. 1945-10: most Japanese armed forces withdrawed from Shanghai

## Measurement

- how population is recorded (not accurate count but by householad then estimates by average)
- why use population density instead of pure population #
- how year variables correspond to the timeline (ex. if an event occured in Nov of 1937, would I take valeus of 1937 as a variable for prior to event occurance or after?)

## Outcome variables
- outcome variable is "population change in International Settlement during 1937-1942" measured in population density

Add graphs, tables and text. Use sub-sub-headings for each outcome variable or update the subheading to be singular.
Some of our data is of penguins (@fig-bills), from @palmerpenguins.

```{r}
#| label: fig-planes
#| fig-cap: Relationship between wing length and width
#| echo: false
#| warning: false
#| message: false

```

## Predictor variables

- most important predictors include time sensitive indicators: Event1 & Event2

Event1: indicator variable for 1937 (1 if year >= 1937 & <= 1941, 0 otherwise)
- mention a bit about the measurement

Event2: indicator variable for 1937 (1 if year >= 1941, 0 otherwise)
- mention a bit about the measurement

# Model

The goal of our modelling strategy is to evaluate the impact of 2 historical events in Shanghai during WWII.

Here we briefly describe the Difference-in-Difference analysis model used to investigate the impact of the Japanese forces taking over the Chinese district and International Settlement in November 1937 and December 1941 on population shift between different Districts in Shanghai.

Background details and diagnostics are included in [Appendix -@sec-model-details].

## Model set-up

Define $y_i$ as the number of seconds that the plane remained aloft. Then $\beta_i$ is the wing width and $\gamma_i$ is the wing length, both measured in millimeters.

\begin{align} 
y_i|\mu_i, \sigma &\sim \mbox{Normal}(\mu_i, \sigma) \\
\mu_i &= \alpha + \beta_i + \gamma_i\\
\alpha &\sim \mbox{Normal}(0, 2.5) \\
\beta &\sim \mbox{Normal}(0, 2.5) \\
\gamma &\sim \mbox{Normal}(0, 2.5) \\
\sigma &\sim \mbox{Exponential}(1)
\end{align}

We run the model in R [@citeR] using the `rstanarm` package of @rstanarm. We use the default priors from `rstanarm`.

### Model justification

We expect a positive relationship between the size of the wings and time spent aloft. In particular...

We can use maths by including latex between dollar signs, for instance $\theta$.

# Results

Our results are summarized in @tbl-modelresults.

```{r}
#| echo: false
#| eval: true
#| warning: false
#| message: false

library(rstanarm)

first_model <-
  readRDS(file = here::here("models/first_model.rds"))
```

```{r}
#| echo: false
#| eval: true
#| label: tbl-modelresults
#| tbl-cap: "Explanatory models of flight time based on wing width and wing length"
#| warning: false

modelsummary::modelsummary(
  list(
    "First model" = first_model
  ),
  statistic = "mad",
  fmt = 2
)
```

# Discussion

## Reason why the French Concession was Exempted from Japanese Occupation
- During WWII, on the European battlefield, France surrendered to Germany on June 22nd of 1940 (Check this & reference with a credible source). Since Japan was on the same side as Germany during that time, Japanese armed forces decided allow the French Concession to retain its own governance.
- However, the French Concession is still constantly under the watch of the Japanese forces that has occupied the rest of Shanghai since 1942. 街上不得有侮辱日本的旗帜或者宣言，官方必须积极配合日本统治者进行有利于日本统治的宣传,etc.(can reference to video summary here)

## Refugees from outside of Shanghai Province
- industrialization was not so popular back in the 1930s, many provinces are still mostly rural. But warfare messed with the land and the yields, causing many farmers to starve and ultimately have to seek refugee (find job) in industrialized city (where income is not dependent on land yield)
- many people flooded to Shanghai for this reason (can insert calculation in data table 20 from reference book here)

### Job Oppurtunities in Shanghai
Aside from the instability caused by the Japanese armed forces, there are a few other reasons for Chinese population to move into the international settlement & French Concession.
1. Job opportunities
- Heart of Shanghai city in the 1930-1940s, high pop density lead to boom of economy = more job opportunities
- many factories are located with the settlements (less so in concession)

### Barrier to Entry in Job Market
2. Barrier to Entry for certain job types
- 旧时候 -- 通讯还没有那么发达的时候，人们都很团结。那个时候同乡会的力量很强大。在外省，从同一个省来的漂泊者会互相帮扶着垄断一个城市的某种行业。例如，在旧上海，要做酒店生意的话你就必须是湖南人，给人做洗脚生意的肯定是苏北人。。。

## Instability after Japanese Occupation ends in WWII: Chinese Civil War
- after the Japanese armed forces withdrawed from Shanghai, what follows is not recovery and rest for the local population
- Competition for power within China's 2 political parties caused full-scale political warfare, now known as the *Chinese Civil War*
- unlike Western political parties, Chinese political parties in the 1900s are more like parties of a throne. They each have their own ideology and an armed forces that follow. Most of all of the provinces are effected by this internal warfare, which ended in 1949 as 国民党 lost and withdrawed to Taiwan.

## Weaknesses and next steps

### Data Measurement Weaknesses
- data was recorded during wartime, many numbers were missing
- base unit for population records is not ppl, it is instead "household". # of ppl in a household is estimated based on historical data
- many people chose to not report or partially report numbers due to convience or economic reasons, data may not be accurate

### Next Steps

\newpage

\appendix

# Appendix {-}

# Additional data details

# Model details {#sec-model-details}

## Posterior predictive check

In @fig-ppcheckandposteriorvsprior-1 we implement a posterior predictive check. This shows...

In @fig-ppcheckandposteriorvsprior-2 we compare the posterior with the prior. This shows...

```{r}
#| eval: true
#| echo: false
#| message: false
#| warning: false
#| label: fig-ppcheckandposteriorvsprior
#| layout-ncol: 2
#| fig-cap: "Examining how the model fits, and is affected by, the data"
#| fig-subcap: ["Posterior prediction check", "Comparing the posterior with the prior"]

pp_check(first_model) +
  theme_classic() +
  theme(legend.position = "bottom")

posterior_vs_prior(first_model) +
  theme_minimal() +
  scale_color_brewer(palette = "Set1") +
  theme(legend.position = "bottom") +
  coord_flip()
```

## Diagnostics

@fig-stanareyouokay-1 is a trace plot. It shows... This suggests...

@fig-stanareyouokay-2 is a Rhat plot. It shows... This suggests...

```{r}
#| echo: false
#| eval: true
#| message: false
#| warning: false
#| label: fig-stanareyouokay
#| fig-cap: "Checking the convergence of the MCMC algorithm"
#| fig-subcap: ["Trace plot", "Rhat plot"]
#| layout-ncol: 2

plot(first_model, "trace")

plot(first_model, "rhat")
```

\newpage

# References