### 2024-12-02 paper/paper.qmd
finished most of the data section, model section, and added to discussion section

---
title: "My title"
subtitle: "Impact Analysis of Japanese Occupation on Population Shifts Across Shanghai’s Districts During WWII"
author: 
  - Sandy Yu
thanks: "Code and data are available at: https://github.com/Jingying-yu/Shanghai_population_change"
date: today
date-format: long
abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
format: pdf
number-sections: true
bibliography: references.bib
---

```{r}
#| include: false
#| warning: false
#| message: false

library(tidyverse)
library(arrow)
library(readr)
library(knitr)
library(kableExtra)
library(modelsummary)
library(rstanarm)
library(dplyr)
library(broom.mixed)
```

```{r}
#| include: false
#| warning: false
#| message: false

calibration_percent <- read_parquet(file = here::here("data/02-analysis_data/calibration_percent.parquet"))
year_refugee_summary <- read_parquet(file = here::here("data/02-analysis_data/year_refugee_summary.parquet"))
combined_data <- read_parquet(file = here::here("data/02-analysis_data/combined_data.parquet"))
model_format <- read_parquet(file = here::here("data/02-analysis_data/model_format.parquet"))
did_model <- readRDS(file = here::here("models/did_model.rds"))
```

# Introduction

Overview paragraph

Estimand paragraph

Results paragraph

Why it matters paragraph

Telegraphing paragraph: The remainder of this paper is structured as follows. @sec-data....

# Data {#sec-data}

## Overview

We use the statistical programming language R [@citeR].... Our data [@shelter].... Following @tellingstories, we consider...

## Data Sources

-   talk about where the data in my primary reference book is gathered
-   mentions the credibility of source and instability of recording instrument & methodology

## Historical Background

Place: Shanghai, China

When: 1936-1942

Who: Chinese population in Shanghai

Define: 1. give 1-2 sentence broad overview of China's state of unrest 2.THREE districts in Shanghai: Chinese District, International Settlement, French Concession - who controlled each district and the level of governance each authority have in comparison to Chinese government 3.Outline area (%) of each district (do not get into specifics, put that in Results section)

Important Event Timeline

1\. 1937-08-13: Japanese armed forces entered Shanghai

2\. 1937-11-12: Japanese armed forces claims occupation of Shanghai --\> ends Chinese district

1940-06-22: France surrendered to Germany --> French Concession stopped taking refugees

3\. 1942-01: Japanese armed forces claims authority over International Settlement (which was mainly under the governance of U.K and U.S prior to this date)

4\. 1945-08-15: Japan surrendered in WWII

5\. 1945-10: most Japanese armed forces withdrew from Shanghai

## Measurement & Methodology

-   how population is recorded
-   Why choose 1936 - 1942 (ex. if an event occurred in Nov of 1937, would I take values of 1937 as a variable for prior to event occurrence or after?)

Survey Difficulties - Population Volatility caused by Warfare: Population is extreme volatile within this time period, population shifts constantly as 战况 proceed - Regional Recording Limitations: Survey response collect difficulty varies between districts. - Resistance due to Political Risk

## Variables of Interest

```{r}
#| echo: false
#| eval: true
#| label: tbl-outcome_variable
#| tbl-cap: "Explanatory models of flight time based on wing width and wing length"
#| warning: false
#| message: false

knitr::kable(combined_data, col.names = c("Year", "Pop CHN District", "Pop IS", "Pop FC"), align = c("c", "c", "c", "c"))
```

### Outcome variable

- outcome variable is *population*

### Predictor variables
Time variable: year
Observations used in modelling and other analysis in this paper is ranged between 1936 to 1942. No observation is selected before 1936 as the growth pattern varies significantly from the event of our interest (Japanese occupation of Shanghai's different districts), and no value is elected beyond 1942 due to data constraints. No annual population records exists on a district bases beyond 1942. Foreign districts 解散 in 1949 as new political party came to power and the People's Republic of China was declared.

#### Binary Indicators
District indicators:
- Variable *district_is* abd *district_fc* are two binary variable used to indicate the district identity of an observation.
There are 3 possible district identity in Shanghai during 1930s-40s: Chinese-administrated District, International Settlement, and French Concession.

The Chinese-administrated District is governed by local Chinese authority. Land type in this area is primary rural, most of the local population are farmers. Percentage of foreign (out-of-province) immigrants or refugees is low. (@sh) In this analysis paper, Chinese-administrated District takes value of *district_is = 0* and *district_fc = 0* (it is neither International Settlement nor French Concession).

The International Settlement district is a generalized term for section of lands that allowed the occupation of foreign (non-Chinese) residence. Foreigners in Shanghai were not allow to reside in areas outside of the bounds of this district. Governance power of foreign population within the district -- including the policing, taxing, etc, are transferred to foreign authorities. This district takes up most of the central-city in Shanghai during the 1930s-40s, with foreign investment attracting many business opportunities and subsequently job opportunities. Due to the complicated political authorities within the district, the part of Shanghai remains initially untouched by the Japanese forces as the entered Shanghai's Chinese-administrated District. International Settlement district takes the value of *district_is = 1* and *district_fc = 0* in this paper.

French Concession district is a land district given solely to the French. French residency are given the authorization to self-govern within this area. Due to political reasons on the Western battlefield, this district remains untouched by the Japanese forces all the way until they withdrew from Shanghai. The French Concesion district takes value of *district_is = 0* and *district_fc = 1* for all analysis in the paper.

Historic Event indicators:
- *cd_occupied*, *french_surrender*, and *is_occupied* are binary variables that mark the whether or not a major historic event that may have impacted Shanghai's population during the observation period (1936-1942) has occured yet.
1. The Japanese armed forces declared occupation of the Chinese-administrated district on 1937-11-12. With expectation that Shanghai's population will internally shift out of the Chinese-administrated district, the binary variable *cd_occupied* is used to measure the impact of this event on the population change in Shanghai.
2. As WWII impacted the Eastern globe, on the primary battlefield (Western Battlefield), France surrendered to Germany on 22 June 1940. Since Japan sided with Germany in WWII, France's whiteflagging to Germany led to Shanghai's Japanese forces promising not to invade the French Concession on the condition that the district stop taking refugees and support Japanese governance. The paper uses the binary variable *french_surrender* to evaluate the impact this event has on the overall population in Shanghai.
3. In January of 1942, the Japanese forces occupied the International Settlement. Local authority handed over its governing power to the invading forces. This event's impact on Shanghai's population is measuring using the binary variable *is_occupied*.

```{r}
#| echo: false
#| eval: true
#| label: tbl-d_variables
#| tbl-cap: "Explanatory models of flight time based on wing width and wing length"
#| warning: false
#| message: false

# Select the rows you want
sliced_table <- model_format |>
  select(year, population, district_is, district_fc, cd_occupied, french_surrender, is_occupied) |>
  slice(c(3, 4, 8, 15, 16, 20))

knitr::kable(sliced_table, col.names = c("Year", "Population", "district_is", "district_fc", "cd_occupied", "french_surrender", "is_occupied"), align = c("c", "c", "c", "c", "c", "c", "c"))

```

# Model

The goal of our modelling strategy is to evaluate the impact of 2 historical events in Shanghai during WWII.

Here we briefly describe the Difference-in-Difference analysis model used to investigate the impact of the Japanese forces taking over the Chinese district and International Settlement in November 1937 and December 1941 on population shift between different Districts in Shanghai.

Background details and diagnostics are included in Appendix \[@sec-model-details\].

## Model set-up

$$y_i|\mu_i, \sigma \sim \text{Normal}(\mu_{i,t}, \sigma)$$

$$\alpha \sim \text{Normal}(0, 2.5)$$

$$\beta_k \sim \text{Normal}(0, 2.5) \quad \text{for } k = 1, ..., 11$$

$$\sigma \sim \text{Exponential}(1)$$

Where:

-   $y_i$: Observed population in district i at year t.
-   $\text{cd_occupied}$: Indicator for whether the Chinese-administered district was occupied by Japanese forces in year t (1 if $t \geq 1937$, 0 otherwise).
-   $\text{french_surrender}$: Indicator for the French Surrendering to Germany in 1940, leading to the French Concession rejecting refugees in Shanghai (1 if $t \geq 1940$, 0 otherwise).
-   $\text{is_occupied}$: Indicator for the Japanese occupation of the International Settlement in 1942 (1 if $t \geq 1942$, 0 otherwise).
-   $\text{district_type}$: Dummy variable for International Settlement districts (1 if $i$ is IS, 0 otherwise).
-   $\text{district_type}$: Dummy variable for French Concession districts (1 if $i$ is FC, 0 otherwise).

Interaction Terms:

The interaction terms in the model are essential because they capture how the impact of historical events (e.g., the 1937 occupation of Chinese-administered districts, the 1940 refugee rejection in the French Concession, and the 1942 occupation of the International Settlement) varies across different district types. By including interactions between event indicators and district types, the model accounts for district-specific responses to each event, controlling for the fact that the effect of an event on population changes may not be uniform across all districts. For instance, the 1937 occupation might lead to population declines in Chinese-administered districts but population increases in the International Settlement due to refugee influx. These terms ensure that the model can differentiate and estimate the unique effects of each event within each district type, thereby improving precision and interpretability.

-   $\text{cd_occupied} \cdot \text{district_type}$: Captures the effect of the 1937 occupation on the International Settlement in comparison to the Chinese-administered district (base district).
-   $\text{cd_occupied} \cdot \text{district_type}$: Captures the effect of the 1937 occupation on the French Concession in comparison to the base district.
-   $\text{french_surrender} \cdot \text{district_type}$: Effect of the 1940 refugee rejection on the International Settlement.
-   $\text{french_surrender} \cdot \text{district_type}$: Effect of the 1940 refugee rejection on the French Concession.
-   $\text{is_occupied} \cdot \text{district_type}$: Effect of the 1942 occupation on the International Settlement.
-   $\text{is_occupied} \cdot \text{district_type}$: Effect of the 1942 occupation on the French Concession.

We run the model in R [@citeR] using the `rstanarm` package of @rstanarm. We use the default priors from `rstanarm`.

### Model justification

We expect a positive relationship between the size of the wings and time spent aloft. In particular...

We can use maths by including latex between dollar signs, for instance $\theta$.

# Results

-   Data Trend Visualization

```{r}
#| label: fig-pop
#| fig-cap: Population in Shanghai between 1936 - 1942
#| echo: false
#| warning: false
#| message: false

# Create the bar chart with a trend line
ggplot(model_format, aes(x = year, y = population, fill = district_type, color = district_type)) +
  geom_bar(stat = "identity", position = "dodge", alpha = 0.7) +  # Bar chart with transparency
  geom_line(aes(group = district_type), size = 1) +  # Add trend line
  geom_point(size = 2) +  # Add points on the trend line
  scale_fill_brewer(palette = "Paired") +  # Use ColorBrewer palette for fill
  scale_color_brewer(palette = "Paired") +  # Use ColorBrewer palette for trend line
  labs(
    title = "Population In Shanghai between 1936-1942",
    x = "Year",
    y = "Population",
    fill = "District Type",
    color = "District Type"
  ) +
  theme_minimal() +
  theme(
    legend.position = "bottom",
    text = element_text(size = 12),
    plot.title = element_text(hjust = 0.5)
  )
```

-   Model results are summarized in @tbl-modelresults in Appendix (@sec-model-details).

```{r}
#| echo: false
#| eval: true
#| warning: false
#| message: false

# Extract coefficients and statistics as a data frame
model_summary <- broom.mixed::tidy(did_model)
```

#### District Impact on Population
```{r}
#| echo: false
#| eval: true
#| label: tbl-district
#| tbl-cap: "Explanatory models of flight time based on wing width and wing length"
#| warning: false

# Filter for the specific coefficient you want (e.g., "cd_occupied")
d_types <- model_summary |>
  filter(term %in% c("(Intercept)", "district_is", "district_fc")) |>
  mutate(
    estimate = round(estimate, 2),
    std.error = round(std.error, 2))

knitr::kable(d_types, col.names = c("Parameters", "Estimated Effect", "Standard Error"), align = c("c", "c", "c"))
```

#### Historical Events
```{r}
#| echo: false
#| eval: true
#| label: tbl-events
#| tbl-cap: "Explanatory models of flight time based on wing width and wing length"
#| warning: false

# Filter for the specific coefficient you want (e.g., "cd_occupied")
event_types <- model_summary |>
  filter(term %in% c("cd_occupied", "french_surrender", "is_occupied")) |>
  mutate(
    estimate = round(estimate, 2),
    std.error = round(std.error, 2))

knitr::kable(event_types, col.names = c("Parameters", "Estimated Effect", "Standard Error"), align = c("c", "c", "c"))
```

#### Historic Events' Impact on Each District
```{r}
#| echo: false
#| eval: true
#| label: tbl-events_district
#| tbl-cap: "Explanatory models of flight time based on wing width and wing length"
#| warning: false

# Filter for the specific coefficient you want (e.g., "cd_occupied")
impact_types <- model_summary[7:12, ] |>
  mutate(estimate = round(estimate, 2),
    std.error = round(std.error, 2))

print(impact_types)
knitr::kable(impact_types, col.names = c("Parameters", "Estimated Effect", "Standard Error"), align = c("c", "c", "c", "c", "c", "c"))
```

# Discussion

## Interpretation of the Event Impacts

### Impact of Occupation in Chinese District

### Impact of French Surrendering to Germany on Shanghai

### Impact of Occupation in International Settlement

## Reason why the French Concession was Exempted from Japanese Occupation

-   During WWII, on the European battlefield, France surrendered to Germany on June 22nd of 1940 (Check this & reference with a credible source). Since Japan was on the same side as Germany during that time, Japanese armed forces decided allow the French Concession to retain its own governance.
-   However, the French Concession is still constantly under the watch of the Japanese forces that has occupied the rest of Shanghai since 1942. 街上不得有侮辱日本的旗帜或者宣言，官方必须积极配合日本统治者进行有利于日本统治的宣传,etc.(can reference to video summary here)

## Refugees from outside of Shanghai Province

-   industrialization was not so popular back in the 1930s, many provinces are still mostly rural. But warfare messed with the land and the yields, causing many farmers to starve and ultimately have to seek refugee (find job) in industrialized city (where income is not dependent on land yield)
-   many people flooded to Shanghai for this reason (can insert calculation in data table 20 from reference book here)

### Job Oppurtunities in Shanghai

Aside from the instability caused by the Japanese armed forces, there are a few other reasons for Chinese population to move into the international settlement & French Concession. 1. Job opportunities - Heart of Shanghai city in the 1930-1940s, high pop density lead to boom of economy = more job opportunities - many factories are located with the settlements (less so in concession)

### Barrier to Entry in Job Market

2.  Barrier to Entry for certain job types

-   旧时候 -- 通讯还没有那么发达的时候，人们都很团结。那个时候同乡会的力量很强大。在外省，从同一个省来的漂泊者会互相帮扶着垄断一个城市的某种行业。例如，在旧上海，要做酒店生意的话你就必须是湖南人，给人做洗脚生意的肯定是苏北人。。。

## Instability after Japanese Occupation ends in WWII: Chinese Civil War

-   after the Japanese armed forces withdrawed from Shanghai, what follows is not recovery and rest for the local population
-   Competition for power within China's 2 political parties caused full-scale political warfare, now known as the *Chinese Civil War*
-   unlike Western political parties, Chinese political parties in the 1900s are more like parties of a throne. They each have their own ideology and an armed forces that follow. Most of all of the provinces are effected by this internal warfare, which ended in 1949 as 国民党 lost and withdrawed to Taiwan.

## Weaknesses and next steps

### Data Measurement Weaknesses

-   data was recorded during wartime, many numbers were missing
-   base unit for population records is not ppl, it is instead "household". \# of ppl in a household is estimated based on historical data
-   many people chose to not report or partially report numbers due to convience or economic reasons, data may not be accurate

### Next Steps

\newpage

\appendix

# Appendix {.unnumbered}

# Additional data details

# Model details {#sec-model-details}

## Model Summary Table
```{r}
#| echo: false
#| eval: true
#| warning: false
#| message: false

library(rstanarm)

did_model <-
  readRDS(file = here::here("models/did_model.rds"))
```

```{r}
#| echo: false
#| eval: true
#| label: tbl-modelresults
#| tbl-cap: "Explanatory models of flight time based on wing width and wing length"
#| warning: false

modelsummary::modelsummary(
  list(
    "Population model" = did_model
  ),
  statistic = "mad",
  fmt = 2,
  metrics = c("R2", "RMSE")
)
```

## Posterior predictive check

In @fig-ppcheckandposteriorvsprior-1 we implement a posterior predictive check. This shows...

In @fig-ppcheckandposteriorvsprior-2 we compare the posterior with the prior. This shows...

```{r}
#| eval: true
#| echo: false
#| message: false
#| warning: false
#| label: fig-ppcheckandposteriorvsprior
#| layout-ncol: 2
#| fig-cap: "Examining how the model fits, and is affected by, the data"
#| fig-subcap: ["Posterior prediction check", "Comparing the posterior with the prior"]

model_format <- read_parquet(file = here::here("data/02-analysis_data/model_format.parquet"))

pp_check(did_model) +
  theme_classic() +
  theme(legend.position = "bottom")

posterior_vs_prior(did_model) +
  theme_minimal() +
  scale_color_brewer(palette = "Set1") +
  theme(legend.position = "bottom") +
  coord_flip()
```

## Diagnostics

@fig-stanareyouokay-1 is a trace plot. It shows... This suggests...

@fig-stanareyouokay-2 is a Rhat plot. It shows... This suggests...

```{r}
#| echo: false
#| eval: true
#| message: false
#| warning: false
#| label: fig-stanareyouokay
#| fig-cap: "Checking the convergence of the MCMC algorithm"
#| fig-subcap: ["Trace plot", "Rhat plot"]
#| layout-ncol: 2

plot(did_model, "trace")

plot(did_model, "rhat")
```

\newpage

# References